Black-hole ringdown as a probe of higher-curvature gravity theories

### 2022-04-25 papers/paper_alt_theor_bounds/paper_alt_theor_bounds.tex
added bbh nr papers, changes to conclusions

\begin{table}[t]
\begin{tabular}{c | c c}
\hline \hline
Theory & Constraint & This work \\
\hline
EdGB        & $\ell_{\rm EdGB} \leqslant 1.18$~km~(GW)~\cite{Lyu:2022gdr} & -- \\
dCS         & $\ell_{\rm dCS} \leqslant 8.5$~km~(EM+GW)~\cite{Silva:2020acr}  & $\ell_{\rm dCS} \leqslant 38.7$~km \\
cubic EFT   & -- & $\ell_{\rm cEFT} \leqslant 38.2$~km \\
quartic EFT & $\ell_{\rm qEFT} \leqslant 150$~km~\cite{Sennett:2019bpc}  & $\ell_{\rm qEFT} \leqslant 51.3$~km\hs{$^{\ast}$} \\
\hline \hline
\end{tabular}
\caption{Summary of the upper limits, at $90\%$ credible level,
on the length-scale of the modified gravity theories under investigation.
They were obtained combining Bayesian-inference results from GW150914 and GW200129.
Current constraints on $\ell_{\rm th}$ are also listed.
%
\hs{Our result for the quartic EFT of GR is only in marginal tension with our hard cut-off scale for the
validity of the theory and hence we do still quote it here. See Sec.~\ref{sec:results_qeft} for details.}
}
\label{tab:bound_summary}
\end{table}

@@

How can we determine the beyond-GR corrections? In GR, comparison between the numerically determined
Kerr QNMs against the fitting formula~\eqref{eq:qnm_nongr_iso} fixes the GR expansion coefficients
$\omega^{(j)}$ and $\tau^{(j)}$.
%
We can proceed in a similar way with QNMs calculated in the context of a non-GR theory.
%
In particular, in the literature, we can already find fitting formulas relating
the QNMs to the BHs mass, spin and coupling constant $\ell_{\rm th}$, the latter being
specific to each theory, up to $j=1$ in the spin expansion (see Table~\ref{tab:ref_theories_qnms}).
%
The idea is then to compare these formulas against Eq.~\eqref{eq:qnm_nongr_iso}
to fix $p$, $\delta\omega^{(j)}$, and $\delta\tau^{(j)}$.
%
Because QNMs of rotating BHs in modified gravity theories are not known to
all spin values, we can expect that the $j=1$ coefficients to be susceptible to
changes as calculations beyond-leading order in spin are accomplished in the future.
%
That is not the case for the $j=0$ coefficients and the situation is the same as in GR,
in which the $j=0$ coefficients are simply the QNMs of the Schwarzschild BH.

@@

There are two conditions that we must verify before we can confidently claim to
have placed a constraint on $\ell_{\rm th}$. First, as we have explained in Sec.~\ref{sec:review_theories}, all theories that we
consider must be interpreted as an EFT, meaning that it should be
considered valid only below an energy scale, or equivalently, a length-scale.
%
As a cut-off length-scale for the validity of the EFT we use,
%
\begin{equation}
\Lambda_{\rm EFT} (\varepsilon, \mathfrak{m}) = \varepsilon \, \frac{G \mathfrak{m}}{c^2} \,,
\label{eq:def_cutoff}
\end{equation}
%
where $\varepsilon$ is a dimensionless number and $\mathfrak{m}$ is the median
value of one of the mass scales involved in the problem.
%
We note that $\Lambda_{\rm EFT}$ has dimensions of length and hence can be compared to the coupling strength $\ell_{\rm th}$.
%
Here we explore the range $\varepsilon \in [0, 1]$, but following
the Refs.~\cite{Nair:2019iur,Perkins:2021mhb,Lyu:2022gdr} we quote our final results using
$\varepsilon = 1/2$.
% %
% In Refs.~\cite{Nair:2019iur,Perkins:2021mhb,Lyu:2022gdr}, $\varepsilon = 1/2$ was used,
% but here we explore $\varepsilon \in [0, 1]$ for generality.
% %
We say that a bound has been placed on $\ell_{\rm th}$, if most of the PDF $P(\ell_{\rm th} | d)$ support is in the interval
$[0, \, \Lambda_{\rm EFT}(1/2, \mathfrak{m})]$.
%
In practice, this can be quantified through the cumulative distribution function
(CDF) associated with the marginalized posterior distribution $P(\ell_{\rm th} | d)$, namely
%
\begin{equation}
P(\ell_{\rm th} \leqslant \ell^{\rm max}_{\rm th} | d) = \int_{0}^{\,\ell^{\rm max}_{\rm th}} \dd \ell'\, P(\ell' | d).
\label{eq:def_cdf}
\end{equation}
%
For instance, we require that for a bound at 90\% credible level to be placed on $\ell_{\rm th}$ that
%
\begin{equation}
% \textrm{CDF}(\ell) \leqslant \textrm{CDF}(T_{\rm EFT}(\varepsilon; \, \mathfrak{m})).
P(\ell_{\rm th} \leqslant \Lambda_{\rm EFT} | d) \leqslant 0.9,
\quad \textrm{(EFT bound)},
\label{eq:eft_bound}
\end{equation}
%
where we let $\ell^{\rm max}_{\rm th} = \Lambda_{\rm EFT}$ in Eq.~\eqref{eq:def_cdf}, and
likewise for other credibility percentiles.

@@

In particular, we found, that within the interpretation of these theories as EFTs and the
region of validity of the ParSpec framework, the fundamental length-scale of dCS gravity is bound as
$\ell_{\rm dCS} \leqslant 34.5$~km, at 90\% credible level, when stacking the posteriors
of GW150914 and GW200129. This is the first GW-alone constrain on this theory.
%
In contrast, we could not place any bounds on the \hs{fundamental} length-scale \abcomm{Maybe
this is common in the literature, but I would call it length-scale instead of coupling
constant. What do you think?}~\hscomm{I agree.} of EdGB gravity $\ell_{\rm EdGB}$.
%
This dichotomy between the two theories has a counterpart with works that considered
the inspiral part of the GW signal alone~\cite{Nair:2019iur,Perkins:2021mhb,Lyu:2022gdr}.
%
Using data of the LVK BBHs, it was found that the posterior distributions for
deviations from GR were uninformative in dCS gravity, but not in EdGB gravity.
We emphasize that both those theories (and the cubic EFT of GR also studied here) all predict the same exponent $p$ in ParSpec.
%
Hence, our results show how the inclusion of theory-specific information into the ParSpec framework
can result in different outcomes for different theories, even if they predict the same value of $p$.

@@

Let us discuss some avenues for future work.
%
First, we could implement a high-spin version of the GR fitting coefficients to the ParSpec formulas.
%
This has already been done in Ref.~\cite{Carullo:2021dui} extending the validity of the ParSpec formulas up to
spins of $\chi_{\rm f} \approx 0.99$.
%
For the events analyzed here, the original fit by Ref.~\cite{Maselli:2019mjd}
was sufficient, but it might not be the case with upcoming GW observation campaigns.
%
Second, it would be important to incorporate additional effects, such as spin-precession and eccentricity
to \pSEOB~(see, for instance, Refs.~\cite{Ossokine:2020kjp,Ramos-Buades:2021adz}). \ab{Third, it will be
interesting, to perfom tests of modified theories of gravity using IMR waveform models that include, during
the inspiral stage, finite-size effects induced by the non-GR geometry around the  holes ---
for example the ones due to spin-induced quadrupole, tidal deformability and absorbtion, and also
orbital effects due to non-GR gravitational interactions
between the holes. Those effects could be included using the flexible theory-independent
method~\cite{Mehta:2022pcn}, as done in Ref.~\cite{Sennett:2019bpc} or the
TIGER code~\cite{Li:2011cg,Agathos:2013upa}. However, setting bounds on deviations
from GR caused by orbital effects requires a different EFT interpretation than what we
adopted in Sec.~\ref{sec:remarks} (see also Sec. II C in Ref.~\cite{Sennett:2019bpc}).
Indeed, in this case one would need to analyze the data considering that the modified theory of
gravity is valid for $\ell \gtrsim  M$, but $\ell \lesssim R$, being $R$ the binary's separation. Fourth,
to test the robustness of the results obtained in this paper, it will be very useful to employ NR waveforms
produced in some of the non-GR theories under consideration, as synthetic signals, and carry out
Bayesian analysis to recover the binary's parameters, including the non-GR ones during the ringdown.
As today, there are only a handful of such NR simulations~XXX. Last, but not least, it will be very
beneficial to calculate the QNMs (complex) frequencies of rapidly-rotating BHs in modified gravity theories
(with BH perturbation theory). This is a challenging problem, but certainly necessary,
also in the context of ParSpec, if higher spin ocrrections would need to be
included to make the framwork robust.}

@@

We start by considering EdGB gravity and focus on
Refs.~\cite{Blazquez-Salcedo:2016enn,Pierini:2021jxd} to determine the ParSpec coefficients for this theory.
% and~\cite{Pierini:2021jxd} for the QNMs of for which the QNMs of nonrotating and
% slowly-rotating BHs where calculated in Refs.~\cite{Pani:2009wy,Blazquez-Salcedo:2016enn,Blazquez-Salcedo:2017txk}
% and~\cite{Pierini:2021jxd}, respectively.
%
In particular, Ref.~\cite{Blazquez-Salcedo:2016enn} found that the damping time
of the \hs{dominant \emph{axial} gravitational-led} mode increases as the coupling constant
$\ell_{\rm EdGB}$ is increased. The leading-order spin corrections to the
\emph{polar}-parity QNMs was studied in Ref.~\cite{Pierini:2021jxd}.
%
Hence, according to the prescription of Sec.~\ref{sec:theory_specific_qnm},
\hs{we select the axial-parity branch of QNMs and
%
for consistency, we only consider the leading-order $j=0$ coefficients in EdGB gravity.}

@@

\hscomm{Could remove this part until the end of this section.}
\hs{We can expand the resulting formula} in $\gamma_{\rm EdGB}$ and the coefficients $\delta \omega^{(i)}$ and $\delta \tau^{(i)}$, $i=1,2$ can be
read-off by comparison against Eqs.~\eqref{eq:qnm_nongr_iso}, where for the damping time we use the relation ${\rm Im}(\sigma)_{\rm EdGB} = - 1 / \tau_{\rm EdGB}$
and re-expand in small coupling and spin.
%
These steps yield \ab{for} $p_{\rm EdGB} = 4$,
%
\begin{align}
     \delta \omega_{\rm EdGB}^{(0)} = -0.0107, \quad \delta \tau_{\rm EdGB}^{(0)} = -0.2480,
\end{align}
%
for the $j=0$ coefficients and
%
\begin{equation}
    \delta \omega_{\rm EdGB}^{(1)} = -0.2480, \quad \delta \tau_{\rm EdGB}^{(1)} = -1.101.
\end{equation}
%
for the $j=1$ coefficients.

@@

For dCS gravity, we follow Ref.~\cite{Wagle:2021tam}, which numerically calculated the QNMs, $\sigma_{\rm dCS}$,
of slowly-rotating BHs, \sout{in this theory} and found that \ab{for} the axial gravitational modes \ab{the} \sout{have their} damping time increases, as we
increase the Chern-Simons coupling $\ell_{\rm dCS}$, at fixed spin $\chi$ of the BH.
%
Hence, according to the prescription of Sec.~\ref{sec:theory_specific_qnm},
this is the branch of QNMs we choose to work with.

We then proceed to determine $\delta\omega^{(j)}$ and $\delta\tau^{(j)}$ as follows.
%
\ab{We use} \sout{Using} the fitting formula Eq.~(54a) of Ref.~\cite{Wagle:2021tam}, namely,
%
\begin{equation}
    M \, {\rm Re}(\sigma)_{\rm dCS} = c_{1} + c_{2} \kappa \zeta + (c_{3} + c_{4} \kappa \zeta) \, (1 - \chi_{\rm f})^{c_{5} + c_{6} \kappa \zeta},
    \label{eq:omega_dsc_eg}
\end{equation}
%
and similarly for the imaginary part, ${\rm Im}(\sigma)_{\rm dCS} =  - 1 / \tau_{\rm dCS}$.
%
Here $\kappa = 1/(16 \pi)$, $\zeta = \ell_{\rm dCS}^{4} / (M^4 \kappa)$, and thus
%
\begin{equation}
    \kappa \zeta = (\ell_{\rm dCS} / M)^4 = \gamma_{\rm dCS},
\end{equation}
%
where $c_{i}$ (with $i=1, \dots, 6$) are fitting coefficients which can be found in
Table II of Ref.~\cite{Wagle:2021tam},

We now expand Eq.~\eqref{eq:omega_dsc_eg} to leading orders in $\chi$ and $\gamma_{\rm dCS}$, and
gather the terms proportional to $\gamma_{\rm dCS}$.
%
We obtain
%
\begin{align}
    M \, \omega_{\rm dCS} =
    \left( 0.3722 + 1.1945 \gamma_{\rm dCS} \right)
    + \left(0.1861 + 5.1828 \gamma_{\rm dCS} \right) \, \chi,
    \nonumber \\
    \label{eq:omega_dcs_interm}
\end{align}
%
where we make use of the numerical values of the coefficients $c_{i}$.
%
We find that (reassuringly) the nonrotating GR part of the expression above agrees with $\omega^{(0)}$ of Ref.~\cite{Maselli:2019mjd} to 0.5\% relative \sout{percent} error. The same \sout{error} estimate \sout{is} \ab{leads to a} larger \ab{relative error} ($\approx 20$\%) for the linear-in-spin coefficient (i.e., 0.1861 in comparison to 0.1258 of Ref.~\cite{Maselli:2019mjd}).
%
We attribute this difference to Ref.~\cite{Wagle:2021tam} having fitted Eq.~\eqref{eq:omega_dsc_eg} to QNM data computed to linear-order in spin, whereas~\cite{Maselli:2019mjd}
fitted Eq.~\eqref{eq:kerr_expansion} to Kerr QNM valid to all orders in spin.

We can now isolate the dCS corrections from Eq.~\eqref{eq:omega_dcs_interm} and
compare against Eq.~\eqref{eq:qnm_nongr_iso}, to find
$p_{\rm dCS} = 4$,
%
\begin{equation}
\delta \omega^{(0)}_{\rm dCS} = 3.1964\,, \quad \delta \omega^{(1)}_{\rm dCS} = 41.199.
\label{eq:cs_omega_coefs}
\end{equation}
%
We can carry the same steps for $\tau_{\rm dCS}$ and find
%
\begin{equation}
\delta \tau^{(0)}_{\rm dCS} = 6.3619\,, \quad \delta \tau^{(1)}_{\rm dCS} = 794.66,
\label{eq:cs_tau_coefs}
\end{equation}
%
which completes the set of fixed non-GR parameters in the ringdown of the \pSEOB{}
waveform model for this theory.
%
As for the case of EdGB gravity, we remark that the alarmingly large values of
$\delta \omega_{\rm dCS}^{(1)}$ and $\delta \tau_{\rm dCS}^{(1)}$ are
compensated by the assumptions that $\gamma_{dCS}$ and $\chi$
are much less than unity, which are the assumptions used in Ref.~\cite{Wagle:2021tam}
to compute the QNMs.

>>>>

\begin{table}[t]
\begin{tabular}{c | c c}
\hline \hline
Theory & Constraint & This work \\
\hline
EdGB        & $\ell_{\rm EdGB} \leqslant 1.18$~km~(GW)~\cite{Lyu:2022gdr} & -- \\
dCS         & $\ell_{\rm dCS} \leqslant 8.5$~km~(EM+GW)~\cite{Silva:2020acr}  & $\ell_{\rm dCS} \leqslant 38.7$~km \\
cubic EFT   & -- & $\ell_{\rm cEFT} \leqslant 38.2$~km \\
quartic EFT & $\ell_{\rm qEFT} \leqslant 150$~km~\cite{Sennett:2019bpc}  & $\ell_{\rm qEFT} \leqslant 51.3$~km\hs{$^{\ast}$} \\
\hline \hline
\end{tabular}
\caption{Summary of the upper limits, at $90\%$ credible level,
on the length-scale of the modified gravity theories under investigation.
They were obtained combining Bayesian-inference results from GW150914 and GW200129.
Current constraints on $\ell_{\rm th}$ are also listed.
%
\hs{Our result for the quartic EFT of GR is only in marginal tension with our hard cut-off scale for the
validity of the theory and hence we do still quote it here. See Secs.~\ref{sec:remarks} and~\ref{sec:results_qeft}
for details.}
}
\label{tab:bound_summary}
\end{table}

@@

How can we determine the beyond-GR corrections? In GR, comparison between the numerically determined
Kerr QNMs against the fitting formula~\eqref{eq:qnm_nongr_iso} fixes the GR expansion coefficients
$\omega^{(j)}$ and $\tau^{(j)}$.
%
We can proceed in a similar way with QNMs calculated in the context of a non-GR theory.
%
In particular, in the literature, we can already find fitting formulas relating
the QNMs to the BHs mass, spin and length-scale $\ell_{\rm th}$, the latter being
specific to each theory, up to $j=1$ in the spin expansion (see Table~\ref{tab:ref_theories_qnms}).
%
The idea is then to compare these formulas against Eq.~\eqref{eq:qnm_nongr_iso}
to fix $p$, $\delta\omega^{(j)}$, and $\delta\tau^{(j)}$.
%
Because QNMs of rotating BHs in modified gravity theories are not known to
all spin values, we can expect that the $j=1$ coefficients to be susceptible to
changes as calculations beyond-leading order in spin are accomplished in the future.
%
That is not the case for the $j=0$ coefficients and the situation is the same as in GR,
in which the $j=0$ coefficients are simply the QNMs of the Schwarzschild BH.

@@

There are two conditions that we must verify before we can confidently claim to
have placed a constraint on $\ell_{\rm th}$. First, as we have explained in Sec.~\ref{sec:review_theories}, all theories that we
consider must be interpreted as an EFT, meaning that it should be
considered valid only below an energy scale, or equivalently, a length-scale.
%
As a cut-off length-scale for the validity of the EFT we use,
%
\begin{equation}
\Lambda_{\rm EFT} (\varepsilon, \mathfrak{m}) = \varepsilon \, \frac{G \mathfrak{m}}{c^2} \,,
\label{eq:def_cutoff}
\end{equation}
%
where $\varepsilon$ is a dimensionless number and $\mathfrak{m}$ is the median
value of one of the mass scales involved in the problem.
%
We note that $\Lambda_{\rm EFT}$ has dimensions of length and hence can be compared to \hs{each theory's fundamental length-scale} $\ell_{\rm th}$.
%
Here we explore the range $\varepsilon \in [0, 1]$, but following
the Refs.~\cite{Nair:2019iur,Perkins:2021mhb,Lyu:2022gdr} we quote our final results using
$\varepsilon = 1/2$.
% %
% In Refs.~\cite{Nair:2019iur,Perkins:2021mhb,Lyu:2022gdr}, $\varepsilon = 1/2$ was used,
% but here we explore $\varepsilon \in [0, 1]$ for generality.
% %
We say that a bound has been placed on $\ell_{\rm th}$, if most of the PDF $P(\ell_{\rm th} | d)$ support is in the interval
$[0, \, \Lambda_{\rm EFT}(1/2, \mathfrak{m})]$.
%
In practice, this can be quantified through the cumulative distribution function
(CDF) associated with the marginalized posterior distribution $P(\ell_{\rm th} | d)$, namely
%
\begin{equation}
P(\ell_{\rm th} \leqslant \ell^{\rm max}_{\rm th} | d) = \int_{0}^{\,\ell^{\rm max}_{\rm th}} \dd \ell'\, P(\ell' | d).
\label{eq:def_cdf}
\end{equation}
%
For instance, we require that for a bound at 90\% credible level to be placed on $\ell_{\rm th}$ that
%
\begin{equation}
% \textrm{CDF}(\ell) \leqslant \textrm{CDF}(T_{\rm EFT}(\varepsilon; \, \mathfrak{m})).
P(\ell_{\rm th} \leqslant \Lambda_{\rm EFT} | d) \leqslant 0.9,
\quad \textrm{(EFT bound)},
\label{eq:eft_bound}
\end{equation}
%
where we let $\ell^{\rm max}_{\rm th} = \Lambda_{\rm EFT}$ in Eq.~\eqref{eq:def_cdf}, and
likewise for other credibility percentiles.

@@

In particular, we found, that within the interpretation of these theories as EFTs and the
region of validity of the ParSpec framework, the fundamental length-scale of dCS gravity is bound as
$\ell_{\rm dCS} \leqslant 34.5$~km, at 90\% credible level, when stacking the posteriors
of GW150914 and GW200129. This is the first GW-alone constrain on this theory.
%
In contrast, we could not place any bounds on the \hs{fundamental} length-scale \abcomm{Maybe
this is common in the literature, but I would call it length-scale instead of coupling
constant. What do you think?}~\hscomm{I agree. I introduced ``length-scale'' at a few other place in the text.} of EdGB gravity $\ell_{\rm EdGB}$.
%
This dichotomy between the two theories has a counterpart with works that considered
the inspiral part of the GW signal alone~\cite{Nair:2019iur,Perkins:2021mhb,Lyu:2022gdr}.
%
Using data of the LVK BBHs, it was found that the posterior distributions for
deviations from GR were uninformative in dCS gravity, but not in EdGB gravity.
We emphasize that both those theories (and the cubic EFT of GR also studied here) all predict the same exponent $p$ in ParSpec.
%
Hence, our results show how the inclusion of theory-specific information into the ParSpec framework
can result in different outcomes for different theories, even if they predict the same value of $p$.

@@

Let us discuss some avenues for future work.
%
First, we could implement a high-spin version of the GR fitting coefficients to the ParSpec formulas.
%
This has already been done in Ref.~\cite{Carullo:2021dui} extending the validity of the ParSpec formulas up to
spins of $\chi_{\rm f} \approx 0.99$.
%
For the events analyzed here, the original fit by Ref.~\cite{Maselli:2019mjd}
was sufficient, but it might not be the case with upcoming GW observation campaigns.
%
Second, it would be important to incorporate additional effects, such as spin-precession and eccentricity
to \pSEOB~(see, for instance, Refs.~\cite{Ossokine:2020kjp,Ramos-Buades:2021adz}).
%
Third, it will be
interesting, to perform tests of modified theories of gravity using IMR waveform models that include, during
the inspiral stage, finite-size effects induced by the non-GR geometry around the BHs ---
for example the ones due to spin-induced quadrupole, tidal deformability and absorption, and also
orbital effects due to non-GR gravitational interactions
between the BHs. Those effects could be included using the flexible theory-independent
method~\cite{Mehta:2022pcn}, as done in Ref.~\cite{Sennett:2019bpc} or the
TIGER code~\cite{Li:2011cg,Agathos:2013upa}. However, setting bounds on deviations
from GR caused by orbital effects requires a different EFT interpretation than what we
adopted in Sec.~\ref{sec:remarks} (see also Sec.~IIC in Ref.~\cite{Sennett:2019bpc}).
Indeed, in this case one would need to analyze the data considering that the modified theory of
gravity is valid for $\ell_{\rm th} \gtrsim  M$, but $\ell_{\rm th} \lesssim \mathcal{R}$, being $\mathcal{R}$ the binary's separation.
%
Fourth, to test the robustness of the results obtained in this paper, it will be very useful to employ NR waveforms
produced in some of the non-GR theories under consideration, as synthetic signals, and carry out
Bayesian analysis to recover the binary's parameters, including the non-GR ones during the ringdown.
As today, there are only a small number of such BBH NR simulations, for a
given theory~\cite{Healy:2011ef,Witek:2018dmd,Okounkova:2017yby,Hirschmann:2017psw,Okounkova:2020rqw,Okounkova:2020rqw,Silva:2020omi,East:2020hgw,East:2021bqk,Figueras:2021abd}.
%
Last, but not least, it will be very
beneficial to calculate the QNMs (complex) frequencies of rapidly-rotating BHs in modified gravity theories
(with BH perturbation theory). This is a challenging problem, but certainly necessary,
also in the context of ParSpec, if higher spin corrections would need to be
included to make the framework robust.

@@

We start by considering EdGB gravity and focus on
Refs.~\cite{Blazquez-Salcedo:2016enn,Pierini:2021jxd} to determine the ParSpec coefficients for this theory.
% and~\cite{Pierini:2021jxd} for the QNMs of for which the QNMs of nonrotating and
% slowly-rotating BHs where calculated in Refs.~\cite{Pani:2009wy,Blazquez-Salcedo:2016enn,Blazquez-Salcedo:2017txk}
% and~\cite{Pierini:2021jxd}, respectively.
%
In particular, Ref.~\cite{Blazquez-Salcedo:2016enn} found that the damping time
of the \hs{dominant \emph{axial} gravitational-led} mode increases as the length-scale
$\ell_{\rm EdGB}$ is increased. The leading-order spin corrections to the
\emph{polar}-parity QNMs was studied in Ref.~\cite{Pierini:2021jxd}.
%
Hence, according to the prescription of Sec.~\ref{sec:theory_specific_qnm},
\hs{we select the axial-parity branch of QNMs and
%
for consistency, we only consider the leading-order $j=0$ coefficients in EdGB gravity.}

@@

\hscomm{Could remove this part until the end of this section.}
\hs{We can expand the resulting formula} in $\gamma_{\rm EdGB}$ and the coefficients $\delta \omega^{(i)}$ and $\delta \tau^{(i)}$, $i=1,2$ can be
read-off by comparison against Eqs.~\eqref{eq:qnm_nongr_iso}, where for the damping time we use the relation ${\rm Im}(\sigma)_{\rm EdGB} = - 1 / \tau_{\rm EdGB}$
and re-expand in $\ell_{\rm EdGB}$ and $\chi$.
%
These steps yield for $p_{\rm EdGB} = 4$,
%
\begin{align}
     \delta \omega_{\rm EdGB}^{(0)} = -0.0107, \quad \delta \tau_{\rm EdGB}^{(0)} = -0.2480,
\end{align}
%
for the $j=0$ coefficients and
%
\begin{equation}
    \delta \omega_{\rm EdGB}^{(1)} = -0.2480, \quad \delta \tau_{\rm EdGB}^{(1)} = -1.101.
\end{equation}
%
for the $j=1$ coefficients.

@@

For dCS gravity, we follow Ref.~\cite{Wagle:2021tam}, which numerically calculated the QNMs, $\sigma_{\rm dCS}$,
of slowly-rotating BHs, and found that for the axial gravitational modes the damping time increases, as we
increase the Chern-Simons coupling $\ell_{\rm dCS}$, \hs{at constant BH spin.}
%
Hence, according to the prescription of Sec.~\ref{sec:theory_specific_qnm},
this is the branch of QNMs we choose to work with.

We then proceed to determine $\delta\omega^{(j)}$ and $\delta\tau^{(j)}$ as follows.
%
Using the fitting formula Eq.~(54a) of Ref.~\cite{Wagle:2021tam}, namely,
%
\begin{equation}
    M {\rm Re}(\sigma)_{\rm dCS} = c_{1} + c_{2} \kappa \zeta + (c_{3} + c_{4} \kappa \zeta) \, (1 - \chi_{\rm f})^{c_{5} + c_{6} \kappa \zeta},
    \label{eq:omega_dsc_eg}
\end{equation}
%
and similarly for the imaginary part, ${\rm Im}(\sigma)_{\rm dCS} =  - 1 / \tau_{\rm dCS}$.
%
Here $\kappa = 1/(16 \pi)$, $\zeta = \ell_{\rm dCS}^{4} / (M^4 \kappa)$, and thus
%
\begin{equation}
    \kappa \zeta = (\ell_{\rm dCS} / M)^4 = \gamma_{\rm dCS},
\end{equation}
%
where $c_{i}$ (with $i=1, \dots, 6$) are fitting coefficients which can be found in
Table II of Ref.~\cite{Wagle:2021tam},

We now expand Eq.~\eqref{eq:omega_dsc_eg} to leading orders in $\chi$ and $\gamma_{\rm dCS}$, and
gather the terms proportional to $\gamma_{\rm dCS}$.
%
We obtain
%
\begin{align}
    M \omega_{\rm dCS} =
    \left( 0.3722 + 1.1945 \gamma_{\rm dCS} \right)
    + \left(0.1861 + 5.1828 \gamma_{\rm dCS} \right) \, \chi,
    \nonumber \\
    \label{eq:omega_dcs_interm}
\end{align}
%
where we make use of the numerical values of the coefficients $c_{i}$.
%
We find (reassuringly) that the nonrotating GR part of the expression above agrees with $\omega^{(0)}$ of Ref.~\cite{Maselli:2019mjd} to 0.5\% relative error. The same estimate leads to a larger relative error ($\approx 20$\%) for the linear-in-spin coefficient (i.e., 0.1861 in comparison to 0.1258 of Ref.~\cite{Maselli:2019mjd}).
%
We attribute this difference to Ref.~\cite{Wagle:2021tam} having fitted
Eq.~\eqref{eq:omega_dsc_eg} to QNM data computed to linear-order in spin,
whereas~\cite{Maselli:2019mjd} fitted Eq.~\eqref{eq:kerr_expansion} to Kerr QNM
valid to all orders in spin.

We can now isolate the dCS corrections from Eq.~\eqref{eq:omega_dcs_interm} and
compare against Eq.~\eqref{eq:qnm_nongr_iso}, to find
$p_{\rm dCS} = 4$,
%
\begin{equation}
\delta \omega^{(0)}_{\rm dCS} = 3.1964\,, \quad \delta \omega^{(1)}_{\rm dCS} = 41.199.
\label{eq:cs_omega_coefs}
\end{equation}
%
We can carry the same steps for $\tau_{\rm dCS} = - 1 / {\rm Im}(\sigma)_{\rm dCS}$ and find
%
\begin{equation}
\delta \tau^{(0)}_{\rm dCS} = 6.3619\,, \quad \delta \tau^{(1)}_{\rm dCS} = 794.66,
\label{eq:cs_tau_coefs}
\end{equation}
%
which completes the set of fixed non-GR parameters in the ringdown of the \pSEOB{}
waveform model for this theory.
%
We remark that the alarmingly large values of
$\delta \omega_{\rm dCS}^{(1)}$ and $\delta \tau_{\rm dCS}^{(1)}$ are
compensated by the assumptions that $\gamma_{\rm dCS}$ and $\chi$
are much less than unity, which are indeed the assumptions used in Ref.~\cite{Wagle:2021tam}
to compute the QNMs.
